---
title: "Lesson 9: Diamonds & Price Predictions"
author: "Josh Goldberg"
date: "October 30, 2017"
output: github_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      warning=FALSE, message=FALSE)
```

```{r Load Libraries}
library(readr)
library(ggplot2)
library(dplyr)
library(tidyr)
library(reshape2)

```
### Welcome
Notes:

***

### Scatterplot Review

```{r Scatterplot Review}
# Let's start by examining two variables in the data set.
# The scatterplot is a powerful tool to help you understand
# the relationship between two continuous variables.

# We can quickly see if the relationship is linear or not.
# In this case, we can use a variety of diamond
# characteristics to help us figure out whether
# the price advertised for any given diamond is 
# reasonable or a rip-off.

# Let's consider the price of a diamond and it's carat weight.
# Create a scatterplot of price (y) vs carat weight (x).

# Limit the x-axis and y-axis to omit the top 1% of values.

data(diamonds)
ggplot(aes(x = carat, y = price),
       data = diamonds) +
  coord_cartesian(xlim = c(0, quantile(diamonds$carat, probs = 0.99)),
  ylim = c(0, quantile(diamonds$price, probs = 0.99))) +
  geom_point(fill = I('#F79420'), color = I('black'), shape = 21)

```

```{r Add a Linear Model}

data(diamonds)
ggplot(aes(x = carat, y = price),
       data = diamonds) +
  coord_cartesian(xlim = c(0, quantile(diamonds$carat, probs = 0.99)),
  ylim = c(0, quantile(diamonds$price, probs = 0.99))) +
  geom_point(color ='#F79420', alpha = 1/4) +
  stat_smooth(method = 'lm')

```

***

### Price and Carat Relationship
Response: As carat weight increases, price increases, possibly exponential. Relationship is non-linear. Dispersion increases with higher carat weight. The linear trend line misses the middle of the data. If lm is used, predictions could be off in certain intervals of the dataset.

***

### Frances Gerety
Notes: 

#### A diamonds is forever.


***

### The Rise of Diamonds
Notes: Diamonds rose through a marketing campaign by De Beers cartel. They controlled the global price of diamonds and convinced consumers of the U.S. to purchase diamonds. https://en.wikipedia.org/wiki/De_Beers#Diamond_monopoly
https://www.theatlantic.com/magazine/archive/1982/02/have-you-ever-tried-to-sell-a-diamond/304575/

***

### ggpairs Function
Notes:

```{r ggpairs Function, message=FALSE, warning=FALSE}
# install these if necessary
install.packages('GGally')
install.packages('scales')
install.packages('memisc')
install.packages('lattice')
install.packages('MASS')
install.packages('car')
install.packages('reshape')
install.packages('plyr')

# load the ggplot graphics package and the others
library(ggplot2)
library(GGally)
library(scales)
library(memisc)

# sample 10,000 diamonds from the data set
set.seed(20022012)
diamond_samp <- diamonds[sample(1:length(diamonds$price), 10000), ]
ggpairs(diamond_samp,
  lower = list(continuous = wrap("points", shape = I('.'))),
  upper = list(combo = wrap("box", outlier.shape = I('.'))),
  axisLabels = 'internal')
```

What are some things you notice in the ggpairs output?
Response: Price is highly correlated with x, y, z, and carat, which are basically the size and dimensions of the diamond. This makes sense. As a diamonds become larger, their price increases.

***

### The Demand of Diamonds
Notes: Distribution of monetary variables tend to highly skewed and vary over orders of magnitude due to path dependence or multiplicative processes (such as inflation).

```{r The Demand of Diamonds}
# Create two histograms of the price variable
# and place them side by side on one output image.

# We’ve put some code below to get you started.

# The first plot should be a histogram of price
# and the second plot should transform
# the price variable using log10.

# Set appropriate bin widths for each plot.
# ggtitle() will add a title to each histogram.

library(gridExtra)

plot1 <- ggplot(aes(x = price), data = diamonds) +
  geom_histogram(aes(fill = I('#099DD9')), binwidth = 100) +
  labs(title = 'Price')

plot2 <- ggplot(aes(x = price), data = diamonds) +
  geom_histogram(aes(fill = I('#F79420')), binwidth = 0.01) +
  scale_x_log10() +
  labs(title = 'Price (log10)')

grid.arrange(plot1, plot2, ncol = 2)

```

***

### Connecting Demand and Price Distributions
Notes: The first histogram has a strong positive skewness due to factors outlined above. Transforming the x scale gives a different prospective. The distribution is now bimodal, which could mean there are two central peaks of pricing (moderate pricing, and high pricing).

***

### Scatterplot Transformation

```{r Scatterplot Transformation}
ggplot(aes(x = carat, y = price), data = diamonds) +
  geom_point() +
  scale_y_log10() +
  labs(title = 'Price Log(10) by Carat')

```


### Create a new function to transform the carat variable

```{r cuberoot transformation}
library(scales)
cuberoot_trans = function() trans_new('cuberoot', transform = function(x) x^(1/3),
                                      inverse = function(x) x^3)
```

#### Use the cuberoot_trans function
```{r Use cuberoot_trans}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point() + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***

### Overplotting Revisited

```{r Sort and Head Tables}
head(sort(table(diamonds$carat), decreasing = T))
head(sort(table(diamonds$price), decreasing = T))

```


```{r Overplotting Revisited}
# Add a layer to adjust the features of the
# scatterplot. Set the transparency to one half,
# the size to three-fourths, and jitter the points.

ggplot(aes(carat, price), data = diamonds) + 
  geom_point(alpha = 1/2, size = 3/4, position = 'jitter') + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***

### Other Qualitative Factors
Notes: Clarity seemed to be a factor, but size is more important. According to Blue Nile, cut has a much bigger impact.

***

### Price vs. Carat and Clarity

Alter the code below.
```{r Price vs. Carat and Clarity}
# install and load the RColorBrewer package
library(RColorBrewer)

ggplot(aes(x = carat, y = price, color = clarity), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
    guide = guide_legend(title = 'Clarity', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
    breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
    breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
```

***

### Clarity and Price
Response: Holding points of the plot constant, it looks that diamonds are almost laways more expensive with higher clarity.

***

### Price vs. Carat and Cut

Alter the code below.
```{r Price vs. Carat and Cut}
library(scales)
ggplot(aes(x = carat, y = price, color = cut), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Cut', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Cut')
```

***

### Cut and Price
Response: Not sure, better cut diamonds appear to be more expensive. Though, there is some overlap.

***

### Price vs. Carat and Color

Alter the code below.
```{r Price vs. Carat and Color}
ggplot(aes(x = carat, y = price, color = color), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Color',
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Color')
```

***

### Color and Price
Response: Color influences price.

***

### Linear Models in R
Notes:

Response:

***

### Building the Linear Model
Notes:

```{r Building the Linear Model}
m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamonds)
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5)
```

Notice how adding cut to our model does not help explain much of the variance
in the price of diamonds. This fits with out exploration earlier.

***

### Model Problems
Video Notes:

Research:
(Take some time to come up with 2-4 problems for the model)
(You should 10-20 min on this)

Response:

***

### A Bigger, Better Data Set
Notes:

```{r A Bigger, Better Data Set}
install.package('bitops')
install.packages('RCurl')
library('bitops')
library('RCurl')

diamondsurl = getBinaryURL("https://raw.github.com/solomonm/diamonds-data/master/BigDiamonds.Rda")
load(rawConnection(diamondsurl))
```

The code used to obtain the data is available here:
https://github.com/solomonm/diamonds-data

## Building a Model Using the Big Diamonds Data Set
Notes:

```{r Building a Model Using the Big Diamonds Data Set}

```


***

## Predictions

Example Diamond from BlueNile:
Round 1.00 Very Good I VS1 $5,601

```{r}
#Be sure you’ve loaded the library memisc and have m5 saved as an object in your workspace.
thisDiamond = data.frame(carat = 1.00, cut = "V.Good",
                         color = "I", clarity="VS1")
modelEstimate = predict(m5, newdata = thisDiamond,
                        interval="prediction", level = .95)
```

Evaluate how well the model predicts the BlueNile diamond's price. Think about the fitted point estimate as well as the 95% CI.

***

## Final Thoughts
Notes:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

